Developmental consistency in children's drawings of object categories

### 2022-07-21 writing/Manuscript_2022.tex
update latest version of manuscript and appendix

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Developmental consistency in children's drawings of object categories},
  pdflang={en-EN},
  pdfkeywords={children's drawings, visual production, tracing, object recognition, visuomotor control},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\begin{center}\begin{threeparttable}}
%   {\end{threeparttable}\end{center}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

@@

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\makeatother
\shorttitle{Development of drawing}
\author{Bria Long\textsuperscript{1}, Ying Wang\textsuperscript{2}, Stella Christie\textsuperscript{2}, Michael C. Frank\textsuperscript{1}, \& Judith E. Fan\textsuperscript{3}}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Stanford University\\\textsuperscript{2} Tsinghua University\\\textsuperscript{3} University of California, San Diego}
\authornote{

Correspondence concerning this article should be addressed to Bria Long, 450 Jane Stanford Way, Stanford CA 94305. E-mail: bria@stanford.edu}
\keywords{children's drawings, visual production, tracing, object recognition, visuomotor control\newline\indent Word count: X}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{csquotes}
\ifxetex
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\else
  \usepackage[shorthands=off,main=english]{babel}
\fi

\title{Developmental consistency in children's drawings of object categories}

\date{}

\abstract{
Childen's drawings of common object categories become dramatically more recognziable across development. What are the major factors that explain this developmental change? Here we examined the degree to which these developmental changes in recognizibility vary across different drawing tasks (i.e.~drawing from observation vs.~from memory), geographical locations (San Jose, US vs.~Beijing, China), and with children's tracing abilities. To do so, we collecting digital drawings of object categories (e.g., cat, airplane) from 4-9 year-olds (N=253). Overall, we show broad consistency in these developmental trajectories across both drawing tasks and these two geographical locations, and we find that children's tracings abilites are good predictors of the recognizability of the drawings that they produce. In addition, we find that our Beijing sample produced more recognizable drawings but showed similar tracing abilities to childen from San Jose, USA. Overall, this work suggests that the developmental trajectory of children's drawings are remarkably consistent and not easily explainable by changes in visuomotor control or working memory capacity.
}

@@

In modern times, drawings are produced prolifically by children from an early age.
Figurative drawings have long provided inspiration for scientists investigating children's emerging cognitive abilities (Minsky \& Papert, 1972), and accordingly a long history of work has examined changes in how children draw objects across development (Fury, Carlson, \& Sroufe, 1997; Karmiloff-Smith, 1990; Kellogg, 1969; Piaget, 1929).
Indeed, there appear to be dramatic changes in how children encode diagnostic visual information in their drawings across age: younger children (4-5 years) tend to include fewer cues in their drawings to differentiate between target concepts (e.g., \textit{adult} vs.~\textit{child}) than older children, who enrich their drawings with more diagnostic part (Sitton \& Light, 1992) and relational (Light \& Simmons, 1983) information.

What drives these dramatic changes in children's drawings across development?
A common view is that these changes are driven primarily by children's increasing ability to plan and control their motor movements (Freeman, 1987; Rehrig \& Stromswold, 2018).
While such changes in visuomotor control are clearly important, this view fails to account for other important constraints on children's drawings --- including what children know about a given target object and how well children are able to access this conceptual knowledge when producing a drawing.

@@

However, this work leaves open the contribution of children's evolving abilities to retrieve and maintain relevant information in memory. Children's ability to access semantic knowledge about categories and to maintain this information in mind when producing a drawing likely changes across childhood (Pailian, Libertus, Feigenson, \& Halberda, 2016). Here, we directly test the idea that a principal reason younger children produce less recognizable drawings is because they simply have more difficulty recalling the relevant visual features of different categories: that is, when asked to \enquote{draw a {[}rabbit{]}}, they may struggle to conjure up the relevant visual details and then hold in mind what rabbits tend to look like. On this account, providing children with additional visual information about different categories -- for example, via canonical photographs of typical exemplars -- could help them improve their drawings of these categories, as it does with adults (Fan, Yamins, \& Turk-Browne, 2018).

@@

To tease apart these alternatives, we investigated the development of children's ability to produce recognizable drawings of visual concepts when children were provided with a verbal cue (\enquote{Can you draw a {[}rabbit{]}?) versus when provided with a picture cue (}Can you draw this {[}rabbit{]} as it looks in the picture?\enquote{). On verbal-cue trials, children thus must access their mental representation of a}rabbit" and choose the features necessary to convey that object's identity. Conversely, on picture-cue trials, children are explicitly asked to rely on the visual features provided in a canonical photograph of each object category.

@@

265 participants were recruited from two local children's museum in {[}BLINDED{]} and at preschool and elementary schools outside of Beijing; approximately equal numbers of participants were recruited in Northern California and the Beijing area. Our goal was to recruit 120 children between 4-9 years of age after exclusions (i.e.~20 4-year-olds, 20 5-year-olds, etc.) at each geographical site. In the San Jose sample, 135 children participated in the experiment; 6 participants were excluded, (3) for skipping more than 6 drawing trials and (3) for scribbling three or more times in a row. Six additional participants were tested but their data was not recorded due to a technical error, and two participants never advanced past the practice trials, leading to a final sample of 121 children. In the Beijing sample, 132 children participated; an additional 8 participants were tested but their data was not recorded due to a technical error with the remote database. Two 10-year-olds (aged 10 years, 0 months and 10 years, 1 month) were accidentally tested and included in the 9-year-old age group. On average, each child contributed 11.46 drawings to analysis (min 6, max = 12). No additional demographic data was recorded about the participants. This protocol was approved by both the Institutional Review Board at {[}blinded{]} (43992, Development of Children's Drawing Abilities) and the Department of Psychology Ethics Committee at {[}blinded{]} in Beijing, China.

@@

Children were seated in front of a touchscreen tablet (Ipad Pro 12.9\enquote{) during each drawing session.
At the beginning of the session, a trained experimenter first told each child, ``After this game is over, someone is going to try to recognize what you were trying to draw. So, please draw so that someone else could guess what you were trying to draw.}
A native English speaker gave these instructions to the San Jose sample, and a native Mandarin speaker gave a translation of these instructions to the Beijing sample.
Following these instructions, children completed 2 shape-tracing trials, providing a baseline measure of visuomotor control in the absence of memory demands.
All children first traced a square, then a more complex shape (see Figure \ref{fig:example-tasks}).
After these tracing trials, children produced drawings of 12 familiar object categories (airplane, bike, bird, car, cat, chair, cup, hat, house, rabbit, tree, watch).
Strokes could not be deleted once drawn, and children had to complete each trial within 30 seconds.
Across trials, we manipulated the type of cue (verbal vs.~picture) children received before producing each drawing (Fig. \ref{fig:example-tasks}), providing a measure of the impact of reminding children of what typical exemplars of each category looked like.
Specifically, children first completed six trials with one cue type, then switched to the other cue type.
The order in which each cue type was used was counterbalanced across children.
On verbal-cue trials, children viewed a short video clip in which an experimenter named the target category: \enquote{What about a {[}cat{]}? Can you draw a {[}cat{]}?}.\\
On picture-cue trials, children viewed a photograph of a typical exemplar of the target category while listening to an audio clip of the same experimenter who said, \enquote{What about this {[}cat{]}? Can you draw the {[}cat{]} as it looks in the picture?}
The photograph remained on the screen for the duration of the drawing trial.
There were three different exemplars of each category used in this study, one of which was randomly sampled on each picture-cue trial.
All experimental code, videos, translations, and stimuli are available on the public repository for this project.

@@

We recorded both the final drawings and the location of each stroke in a digital format, allowing us to precisely measure several different variables that provided proxies for the amount of effort children invested in producing each drawing.
Specifically, we measured: (1) the amount of time spent (i.e., end time of last stroke --- start time of first stroke), (2) the number of individual strokes drawn, and (3) the amount of \enquote{ink} used (i.e., proportion of drawing canvas that was filled).

@@

As in prior work (Long et al., 2021), we used a semi-automated procedure for evaluating how accurately each child performed the tracing task that was validated against empirical judgments of tracing quality. In brief, we decomposed tracing accuracy into two components: a shape error component and a spatial error component. Shape error reflects how closely the participant's tracing matched the contours of the target shape; the spatial error reflects how closely the location, size, and orientation of the participant's tracing matched the target shape. These two error components were computed automatically for each tracing and used to yield a \enquote{tracing score} for each tracing that mirrors adult human judgments of tracing quality (see Appendix for details).

@@

We measured the recognizability of each drawing via an online recognition experiment.
Adult participants based in the U.S. were recruited via Prolific for a 15-minute experiment, compensated at \$14/hour, and asked to identify the category depicted in a random subset of approximately 140 drawings.
Each subset of drawings was balanced with respect to age, category, and site, and each drawing was shown to 10 participants.
Participants were shown these drawings in a random sequence and asked \enquote{What does this look like?}
Participants selected their responses from the set of 12 categories and were encouraged to provide their best guess if they were unsure.
No participants were excluded from analysis for missing the catch trial, which was included to verify that participants could accurately describe their goal in this task.
We then computed a recognition score for each drawing, reflecting the proportion of participants who correctly identified the target category.

@@

In a secondary analysis, we investigated the degree to which any of the above effects were mediated by children's tracing abilities or the amount of effort that children expended while drawings.
We fit the same model above with additional fixed-effects terms for each child's estimated tracing score (see Measuring Tracing Abilities), the time that child spent drawing (in seconds), the amount of \enquote{ink} used (i.e.~percentage of non-white pixels in the drawn image), and the number of strokes produced.
These additional predictors were first z-scored before being included in the model.
Finally, we also assessed the degree to which tracing ability development differed across geographical sites, where tracing scores were modeled as a function of age (in years), site, and their interaction, with the same random-effects structure as the first model.

@@

We found steady changes in the recognizability of children's drawings as a function of age (Figure \ref{fig:main-results}), replicating prior work using the verbal-cue paradigm under less controlled settings (Long et al., 2021).
Moreover, we found that this age-related change did not depend on the drawing task children completed: Children's drawings were just as recognizable when they were cued with a photograph (\enquote{Draw this rabbit as it looks in the picture}) as when they were cued with the category label (\enquote{Can you draw a rabbit?}).\\
While we thought that older children might produce more recognizable drawings when cued with a photograph of a highly prototypical exemplar, as adult participants do (Yang \& Fan, 2021), we did not find evidence of an interaction between age and cue type.
However, we did observe a main effect of geographical site: children in Beijing, China produced drawings that were more recognizable than did children in San Jose, USA (Table 1), reflecting steeper age-related gains earlier in development.

To follow up on this effect of site, we tested two potential sources of variation: visuomotor control and effort.
Insofar as differences in visuomotor control could explain differences between groups, we reasoned that this would manifest as a difference in tracing ability that predicts variation in drawing recognizability.
Instead, we found that children across the two sites were indistinguishable with respect to tracing ability on average (Figure \ref{fig:tracing}), although individual tracing scores were strong predictors of the recognizablity of their own drawings, as in prior work (Long et al., 2021).
Next, we evaluated whether there were systematic differences in effort between the two groups, as measured by the amount of time each child spent producing their drawings and the number of strokes drawn.
Figure \ref{fig:effort-covariates} shows three effort covariates -- average intensity, number of strokes used, and time spent drawing -- measured for each drawing as a function of children's age, drawing task, and geographical site.\\
We found that children spent more time drawing when prompted with a picture cue than a verbal cue, and that children in the Beijing group spent more time on their drawings than their counterparts in San Jose.

@@

Here we examined developmental change in children's ability to produce recognizable drawings of familiar visual concepts, and the degree to which this ability is constrained by retrieval of diagnostic attributes from memory.
We manipulated memory demands by either cueing children with a category label (\enquote{Can you draw a {[}cat{]}?}) or a color photo (\enquote{Can you draw this {[}cat{]} as it looks in the picture?}), and measured the recognizability of each drawing to adult observers.
To evaluate the generality of our findings, we recruited two groups of children in different sites: San Jose, USA and Beijing, China.
We found steady improvement from 4-9 years of age in both groups of children, replicating prior work conducted in a field setting (Long et al., 2021).
We did not find an effect of cue type, however, suggesting that younger children's drawings are unlikely to be constrained by their ability to recall category-diagnostic visual features when they are trying to draw them.

We also discovered that drawing recognizability differed between sites, with children in the Beijing group having produced more recognizable drawings than those in the San Jose group. In follow-up analyses, we found that while there was substantial individual variation in tracing ability and effort, these covariates were insufficient to account for differences in drawing recognizability between groups.
We also discovered that the groups differed with respect to which visual concepts were easier for children to convey in their drawings, perhaps reflecting different patterns of exposure to these concepts.

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  man]{apa6}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Developmental consistency in children's drawings of object categories},
  pdfauthor={Bria Long1, Ying Wang2, Stella Christie2, Michael C. Frank1, \& Judith E. Fan3},
  pdflang={en-EN},
  pdfkeywords={children's drawings, visual production, tracing, object recognition, visuomotor control},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\centering\begin{threeparttable}}
%   {\end{threeparttable}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}

@@

% Overwrite redefinition of paragraph and subparagraph by the default LaTeX template
% See https://github.com/crsh/papaja/issues/292
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\itshape\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-\z@\relax}%
  {\normalfont\normalsize\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{children's drawings, visual production, tracing, object recognition, visuomotor control\newline\indent Word count: X}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{csquotes}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Developmental consistency in children's drawings of object categories}
\author{Bria Long\textsuperscript{1}, Ying Wang\textsuperscript{2}, Stella Christie\textsuperscript{2}, Michael C. Frank\textsuperscript{1}, \& Judith E. Fan\textsuperscript{3}}
\date{}

\shorttitle{Development of drawing}

\authornote{

Correspondence concerning this article should be addressed to Bria Long, 450 Jane Stanford Way, Stanford CA 94305. E-mail: \href{mailto:bria@stanford.edu}{\nolinkurl{bria@stanford.edu}}

\affiliation{\vspace{0.5cm}\textsuperscript{1} Stanford University\\\textsuperscript{2} Tsinghua University\\\textsuperscript{3} University of California, San Diego}

\abstract{%
Children's drawings of common object categories become dramatically more recognziable across development. What are the major factors that explain this developmental change? Here we examined the degree to which these developmental changes in recognizability vary across different drawing tasks (i.e.~drawing from observation vs.~from memory), geographical locations (San Jose, US vs.~Beijing, China), and with children's tracing abilities. To do so, we collected digital drawings of object categories (e.g., cat, airplane) from 4-9 year-olds (N=253). We found that the developmental trajectory of drawing recognizability was remarkably similar when children were asked to draw from observation vs.~memory and across these two geographical locations. In addition, we found that our Beijing sample produced more recognizable drawings but showed similar tracing abilities to children from San Jose, USA. Overall, this work suggests that the developmental trajectory of children's drawings is remarkably consistent and not easily explainable by changes in domain general changes in visuomotor control or working memory.
}

@@

In modern times, drawings are produced prolifically by children from an early age.
Figurative drawings have long provided inspiration for scientists investigating children's emerging cognitive abilities (Minsky \& Papert, 1972), and accordingly a long history of work has examined changes in how children draw objects across development (Fury, Carlson, \& Sroufe, 1997; Karmiloff-Smith, 1990; Kellogg, 1969; Piaget, 1929).
Indeed, there appear to be dramatic changes in how children encode diagnostic visual information in their drawings across age: younger children (4-5 years) tend to include fewer cues in their drawings to differentiate between categories (e.g., \textit{adult} vs.~\textit{child}) than older children, who enrich their drawings with more diagnostic part (Sitton \& Light, 1992) and relational (Light \& Simmons, 1983) information.

What drives these dramatic changes in children's drawings across development?
A common view is that these changes are driven primarily by children's increasing ability to plan and control their motor movements (Freeman, 1987; Rehrig \& Stromswold, 2018).
While such changes in visuomotor control are clearly important, this view fails to account for other important constraints on children's drawings --- including what children know about a given visual concept and how well children are able to access this conceptual knowledge when producing a drawing.

@@

However, this work leaves open the contribution of children's evolving abilities to retrieve and maintain relevant information in memory. Children's ability to access semantic knowledge about categories and to maintain this information in mind when producing a drawing likely changes across childhood (Pailian, Libertus, Feigenson, \& Halberda, 2016). Here, we directly test the idea that a principal reason younger children produce less recognizable drawings is because they simply have more difficulty recalling the relevant visual features of different categories: that is, when asked to ``draw a {[}rabbit{]}'', they may struggle to conjure up the relevant visual details and then hold in mind what rabbits tend to look like. On this account, providing children with additional visual information about different categories -- for example, via canonical photographs of typical exemplars -- could help them improve their drawings of these categories, as it does with adults (Fan, Yamins, \& Turk-Browne, 2018).

@@

To tease apart these alternatives, we investigated the development of children's ability to produce recognizable drawings of visual concepts when children were provided with a verbal cue (``Can you draw a {[}rabbit{]}?) versus when provided with a picture cue (''Can you draw this {[}rabbit{]} as it looks in the picture?``). On verbal-cue trials, children thus must access their mental representation of a''rabbit'' and choose the features necessary to convey that object's identity. Conversely, on picture-cue trials, children are explicitly asked to rely on the visual features provided in a canonical photograph of each object category.

@@

265 participants were recruited from two local children's museum in San Jose and at preschool and elementary schools outside of Beijing; approximately equal numbers of participants were recruited in Northern California and the Beijing area. Our goal was to recruit 120 children between 4-9 years of age after exclusions (i.e.~20 4-year-olds, 20 5-year-olds, etc.) at each geographical site. In the San Jose sample, 135 children participated in the experiment; 6 participants were excluded, (3) for skipping more than 6 drawing trials and (3) for scribbling three or more times in a row. Six additional participants were tested but their data was not recorded due to a technical error, and two participants never advanced past the practice trials, leading to a final sample of 132 children. In the Beijing sample, 121 children participated; an additional 8 participants were tested but their data was not recorded due to a technical error with the remote database. Two 10-year-olds (aged 10 years, 0 months and 10 years, 1 month) were accidentally tested and included in the 9-year-old age group. On average, each child contributed 11.46 drawings to analysis (min 6, max = 12). No additional demographic data was recorded about the participants. This protocol was approved by both the Institutional Review Board at {[}blinded{]} (43992, Development of Children's Drawing Abilities) and the Department of Psychology Ethics Committee at {[}blinded{]} in Beijing, China.

@@

Children were seated in front of a touchscreen tablet (Ipad Pro 12.9'') during each drawing session.
At the beginning of the session, a trained experimenter first told each child, ``After this game is over, someone is going to try to recognize what you were trying to draw. So, please draw so that someone else could guess what you were trying to draw.''
A native English speaker gave these instructions to the San Jose sample, and a native Mandarin speaker gave a translation of these instructions to the Beijing sample.
Following these instructions, children completed 2 shape-tracing trials, providing a baseline measure of visuomotor control in the absence of memory demands.
All children first traced a square, then a more complex shape (see Figure \ref{fig:example-tasks}).
After these tracing trials, children produced drawings of 12 familiar object categories (airplane, bike, bird, car, cat, chair, cup, hat, house, rabbit, tree, watch).
Strokes could not be deleted once drawn, and children had to complete each trial within 30 seconds.
Across trials, we manipulated the type of cue (verbal vs.~picture) children received before producing each drawing (Fig. \ref{fig:example-tasks}), providing a measure of the impact of reminding children of what typical exemplars of each category looked like.
Specifically, children first completed six trials with one cue type, then switched to the other cue type.
The order in which each cue type was used was counterbalanced across children.
On verbal-cue trials, children viewed a short video clip in which an experimenter named the target category: ``What about a {[}cat{]}? Can you draw a {[}cat{]}?''.\\
On picture-cue trials, children viewed a photograph of a typical exemplar of the target category while listening to an audio clip of the same experimenter who said, ``What about this {[}cat{]}? Can you draw the {[}cat{]} as it looks in the picture?''
The photograph remained on the screen for the duration of the drawing trial.
There were three different exemplars of each category used in this study, one of which was randomly sampled on each picture-cue trial.
All experimental code, videos, translations, and stimuli are available on the public repository for this project.

@@

We recorded both the final drawings and the location of each stroke in a digital format, allowing us to precisely measure several different variables that provided proxies for the amount of effort children invested in producing each drawing.
Specifically, we measured: (1) the amount of time spent (i.e., end time of last stroke --- start time of first stroke), (2) the number of individual strokes drawn, and (3) the amount of ``ink'' used (i.e., proportion of drawing canvas that was filled).

@@

As in prior work (Long et al., 2021), we used a semi-automated procedure for evaluating how accurately each child performed the tracing task that was validated against empirical judgments of tracing quality. In brief, we decomposed tracing accuracy into two components: a shape error component and a spatial error component. Shape error reflects how closely the participant's tracing matched the contours of the target shape; the spatial error reflects how closely the location, size, and orientation of the participant's tracing matched the target shape. These two error components were computed automatically for each tracing and used to yield a ``tracing score'' for each tracing that mirrors adult human judgments of tracing quality (see Appendix for details).

@@

We measured the recognizability of each drawing via an online recognition experiment.
Adult participants based in the U.S. were recruited via Prolific for a 15-minute experiment, compensated at \$14/hour, and asked to identify the category depicted in a random subset of approximately 140 drawings.
Each subset of drawings was balanced with respect to age, category, and site, and each drawing was shown to 10 participants.
Participants were shown these drawings in a random sequence and asked ``What does this look like?''
Participants selected their responses from the set of 12 categories and were encouraged to provide their best guess if they were unsure.
No participants were excluded from analysis for missing the catch trial, which was included to verify that participants could accurately describe their goal in this task.
We then computed a recognition score for each drawing, reflecting the proportion of participants who correctly identified the target category.

@@

In a secondary analysis, we investigated the degree to which any of the above effects were mediated by children's tracing abilities or the amount of effort that children expended while drawings.
We fit the same model above with additional fixed-effects terms for each child's estimated tracing score (see Measuring Tracing Abilities), the time that child spent drawing (in seconds), the amount of ``ink'' used (i.e.~percentage of non-white pixels in the drawn image), and the number of strokes produced.
These additional predictors were first z-scored before being included in the model.
Finally, we also assessed the degree to which tracing ability development differed across geographical sites, where tracing scores were modeled as a function of age (in years), site, and their interaction, with the same random-effects structure as the first model.

@@

We found steady changes in the recognizability of children's drawings as a function of age (Figure \ref{fig:main-results}), replicating prior work using the verbal-cue paradigm under less controlled settings (Long et al., 2021).
Moreover, we found that this age-related change did not depend on the drawing task children completed: Children's drawings were just as recognizable when they were cued with a photograph (``Draw this rabbit as it looks in the picture'') as when they were cued with the category label (``Can you draw a rabbit?'').\\
While we thought that older children might produce more recognizable drawings when cued with a photograph of a highly prototypical exemplar, as adult participants do (Yang \& Fan, 2021), we did not find evidence of an interaction between age and cue type.
However, we did observe a main effect of geographical site: children in Beijing, China produced drawings that were more recognizable than did children in San Jose, USA (Table 1), reflecting steeper age-related gains earlier in development.

To follow up on this effect of site, we tested two potential sources of variation: visuomotor control and effort.
Insofar as differences in visuomotor control could explain differences between groups, we reasoned that this would manifest as a difference in tracing ability that predicts variation in drawing recognizability.
Instead, we found that children across the two sites were indistinguishable with respect to tracing ability on average (Figure \ref{fig:tracing}), although individual tracing scores were strong predictors of the recognizability of their own drawings, as in prior work (Long et al., 2021).
Next, we evaluated whether there were systematic differences in effort between the two groups, as measured by the amount of time each child spent producing their drawings and the number of strokes drawn.
Figure \ref{fig:effort-covariates} shows three effort covariates -- average intensity, number of strokes used, and time spent drawing -- measured for each drawing as a function of children's age, drawing task, and geographical site.\\
We found that children spent more time drawing when prompted with a picture cue than a verbal cue, and that children in the Beijing group spent more time on their drawings than their counterparts in San Jose.

@@

Here we examined developmental change in children's ability to produce recognizable drawings of familiar visual concepts, and the degree to which this ability is constrained by retrieval of diagnostic attributes from memory.
We manipulated memory demands by either cueing children with a category label (``Can you draw a {[}cat{]}?'') or a color photo (``Can you draw this {[}cat{]} as it looks in the picture?''), and measured the recognizability of each drawing to adult observers.
To evaluate the generality of our findings, we recruited two groups of children in different sites: San Jose, USA and Beijing, China.
We found steady improvement from 4-9 years of age in both groups of children, replicating prior work conducted in a field setting (Long et al., 2021). We did not find an effect of drawing cue type, however, suggesting that younger children's drawings are unlikely to be constrained by their ability to recall category-diagnostic visual features when they are trying to draw them.

We also discovered that drawing recognizability differed between sites: children in the Beijing group produced more recognizable drawings than those in the San Jose group. While these group-level differences are intriguing, they remain consistent with wide variety of potential explanations.One idea is that children in the Beijing group drew at their preschool and elementary schools, whereas children in the San Jose group drew in a room at a children's science museum; these two testing environments could have created motivational differences that affected how much effort children put into their drawings. However, contra a strong version of this account, we found that the substantial individual variation in basic shape tracing ability and effort (as assessed via ink, time, and strokes spent while drwaing) was insufficient to account for differences in drawing recognizability between groups.
